---
title: "Homework 5"
author: "Rich Chu"
output: html_notebook
---

# Homework 5: To Err is Human; to Rank, Divine

## Part 0

### Problem 0

https://github.com/RichChu1/ma350-hw5

## Part I: Think

### Problem 1

```{r}
rankcor <-function(x, y){  
  n <-length(x)             # Gets the length of x
  stopifnot(length(y)==n)   # Stops if x and y are not the same length
  x.ranks <-rank(x)         # Gets the ranks of x
  y.ranks <-rank(y)         # Gets the ranks of y
  
  mean.x <-mean(x.ranks)    # Gets the mean of the ranks of x
  mean.y <-mean(y.ranks)    # Gets the mean of the ranks of y
  
  covariance.term <-cov(x.ranks - mean.x, y - mean.y)  # Gets the covariance of means of the ranks of x and y
  
  sd.x <-sd(x.ranks)  # Gets the standard deviation of the ranks of x
  sd.y <-sd(y.ranks)  # Gets the standard deviation of the ranks of y
  
  rank.cor <- covariance.term/(sd.x*sd.y)   # Divides the variance of the means of the ranks of x and y by the product of the standard deviation of the ranks of x and y
  
  return(rank.cor)    # Returns the correlation between the standard deviation of the ranks of x and y
  }   

?cov
```

### Problem 2

```{r}
x <- c(2, 4, 6, 8)
y <- c(64, 36, 16, 4)

rankcor(x,y)
```
The rank correlation is -20

### Problem 3

The answer is wrong because the answer must be between -1 and 1 for the problem where: 
1 if there is a perfect, positive linear association
-1 if there is a perfect, negative linear association

### Problem 4

```{r}
x2 <- c(1, 2, 3, 4)
y2 <- c(1, 2, 3, 4)

rankcor(x2,y2)
```
I used 1,2,3,4 as my x and y. This works because a set has the same as itself, so they have a positive linear association.

## Part II: Test

### Problem 5

```{r}
test.rankcor <- function(fun=rankcor){
  x <- c(2, 4, 6, 8)
  y <- c(64, 36, 16, 4)
  
  x2 <- c(1, 2, 3, 4)
  y2 <- c(1, 2, 3, 4)
  
  test <- vector(length= 2)
  
  test[1] <- (as.integer(fun(x, y)) >= -1)  && (as.integer(fun(x, y)) <= 1)

  test[2] <- (as.integer(fun(x2, y2)) >= -1) && (as.integer(fun(x2, y2)) <= 1)
  
  if(test[1] == TRUE && test[2] == TRUE){
    return(TRUE)
  }
  else{
    r <- c("Test 1 is correct","Test 2 is correct")
    if(test[1] == FALSE){
      r[1] <- "Test 1 is incorrect"
    }
    if(test[2] == FALSE){
      r[2] <- "Test 2 is incorrect"
    }
  }
  return(r)
} 
test.rankcor()
```

### Problem 6

The calculation of covariance.term requires the ranks of x minus the mean of the ranks of x and the ranks of y minus the mean of the ranks of y, but in the fuction it is not calculating the ranks of y instead it is using just y. This will make any vector that is not in rank order(least to greatest) will receive an error when runnning

### Problem 7

```{r}
rankcor <-function(x, y){  
  n <-length(x)             
  stopifnot(length(y)==n)   
  x.ranks <-rank(x)         
  y.ranks <-rank(y)        
  
  mean.x <-mean(x.ranks)
  mean.y <-mean(y.ranks)
  
  # covariance.term <-cov(x.ranks - mean.x, y - mean.y)
  covariance.term <-cov(x.ranks - mean.x, y.ranks - mean.y) # This is the change to fix the bug, replace y                                                                with y.ranks
  sd.x <-sd(x.ranks)  
  sd.y <-sd(y.ranks)
  
  rank.cor <- covariance.term/(sd.x*sd.y)  
  
  return(rank.cor)    
}   

test.rankcor()
```

## Part III: Think Some More

### Problem 8

The average rank of a vector with n observations is  always n/2 + 1/2 because the vector of ranks will always be a consecutive list containing all integers within the upper and lower bounds of the vector, making the mean be the actual middle value if n is odd and the middle of the middle 2 values if n is even.

```{r}
rankcor <-function(x, y){  
  n <-length(x)             
  stopifnot(length(y)==n)   
  x.ranks <-rank(x)         
  y.ranks <-rank(y)        
  
  if(n %% 2 == 0){                    
    mean.x <- (x.ranks[n/2] + 1/2)
    mean.y <- (y.ranks[n/2] + 1/2)
  }
  else{
    mean.x <- x.ranks[(n/2) + 1/2]
    mean.y <- y.ranks[(n/2) + 1/2]
  }
  covariance.term <-cov(x.ranks - mean.x, y.ranks - mean.y)
  
  sd.x <-sd(x.ranks)  
  sd.y <-sd(y.ranks)
  
  rank.cor <- covariance.term/(sd.x*sd.y)  
  
  return(rank.cor)    
}   

test.rankcor()
```

### Problem 9
There is no reason to subtract off the mean rank before calculating the sample covariance because the value will always equal 0

```{r}
rankcor <-function(x, y){  
  n <-length(x)             
  stopifnot(length(y)==n)   
  x.ranks <-rank(x)         
  y.ranks <-rank(y)        
  
  covariance.term <-cov(x.ranks, y.ranks)   # Removed the subraction of then mean
  
  sd.x <-sd(x.ranks)  
  sd.y <-sd(y.ranks)
  
  rank.cor <- covariance.term/(sd.x*sd.y)  
  
  return(rank.cor)    
}   

test.rankcor()
```

### Problem 10

```{r}
rankcor <-function(x, y){  
  n <-length(x)             
  stopifnot(length(y)==n)   
  x.ranks <-rank(x)         
  y.ranks <-rank(y)        
  
  covariance.term <-cov(x.ranks, y.ranks)
  
  sd.x <-sqrt(var(x.ranks))  
  sd.y <-sqrt(var(y.ranks))
  
  rank.cor <- covariance.term/(sd.x*sd.y)  
  
  return(rank.cor)    
}   

test.rankcor()
```
