---
title: "Homework5"
author: "Rich Chu"
output: html_notebook
---

# Homework 5: To Err is Human; to Rank, Divine

## Part 0

### Problem 0

https://github.com/RichChu1/ma350-hw5

## Part I

### Problem 1

```{r}
rankcor <-function(x, y){  
  n <-length(x)             # Gets the length of x
  stopifnot(length(y)==n)   # Stops if x and y are not the same length
  x.ranks <-rank(x)         # Gets the ranks of x
  y.ranks <-rank(y)         # Gets the ranks of y
  
  mean.x <-mean(x.ranks)    # Gets the mean of the ranks of x
  mean.y <-mean(y.ranks)    # Gets the mean of the ranks of y
  
  covariance.term <-cov(x.ranks - mean.x, y - mean.y)  # Gets the covariance of means of the ranks of x and y
  
  sd.x <-sd(x.ranks)  # Gets the standard deviation of the ranks of x
  sd.y <-sd(y.ranks)  # Gets the standard deviation of the ranks of y
  
  rank.cor <- covariance.term/(sd.x*sd.y)   # Divides the variance of the means of the ranks of x and y by the product of the standard deviation of the ranks of x and y
  
  return(rank.cor)}   # Returns the correlation between the standard deviation of the ranks of x and y

?cov
```

### Problem 2

```{r}
x <- c(2, 4, 6, 8)
y <- c(64, 36, 16, 4)

rankcor(x,y)
```
The rank correlation is -20

### Problem 3

The answer is wrong because the answer must be between -1 and 1 for the problem where: 
1 if there is a perfect, positive linear association
-1 if there is a perfect, negative linear association

### Problem 4

```{r}
x2 <- c(1, 2, 3, 4)
y2 <- c(1, 2, 3, 4)

rankcor(x2,y2)
```
I used 1,2,3,4 as my x and y. This works because a set has the same as itself, so they have a positive linear association.

## Part II

### Problem 5



### Problem 6



### Problem 7



## Part III

### Problem 8



### Problem 9



### Problem 10




